Don't Mind the Gap: Bridging Network-wide Objectives and Device-level Configurations

### 2016-01-30 paper/compilation.tex
final edits

     pol     &::=& p_1, \dots, p_n & \textit{constraints} \\
     p       &::=& t \hspace{.3em} \Path \hspace{.3em} r_1 \Prefer \dots \Prefer r_m & \textit{preferences} \\
     x       &::=& a.b.c.d/n & \textit{prefix} \\
     c       &::=& \cdots & \textit{community value} \\
     t       &::=& \True & \textit{true} \\
         &\BNFALT& \NOT t & \textit{negation} \\
         &\BNFALT& t_1 \OR{} t_2 & \textit{disjunction} \\
         &\BNFALT& t_1 \AND{} t_2 & \textit{conjunction} \\
         &\BNFALT& \textit{prefix} = x & \textit{prefix test} \\
         &\BNFALT& \textit{comm} = c & \textit{community test} \\
     r       &::=& n & \textit{AS number} \\
         &\BNFALT& \In & \textit{internal loc} \\
         &\BNFALT& \Out & \textit{external loc} \\
         &\BNFALT& r_1 \cup r_2 & \textit{union} \\
         &\BNFALT& r_1 \cap r_2 & \textit{intersection} \\
         &\BNFALT& r_1 \cdot r_2 & \textit{concatenation} \\
         &\BNFALT& \NOT(r) & \textit{path negation} \\
         &\BNFALT& r^* & \textit{repetition} \\
     l       &::=& r_1 \rightarrow r_2 & \textit{link pairs} \\
     cc     &::=& agg(x, l) \BNFALT tag(c, t, l) & \textit{control constraints} \\
  \end{array} \]

  \end{minipage}
  %
  ~~
  \vrule
  ~~
  %
  \begin{minipage}[t]{.5\linewidth}\small
  \hdr{\large Propane Expansions}{}
  \vspace*{-1\baselineskip}
  %
  \[\begin{array}{rcl}
    \hline
    \Any           & = & \Out^* \cdot \In^+ \cdot \Out^* \\
    \Internal      & = & \In^+ \\
    \Only(X)       & = & \Any \cap X^* \\
    \Never(X)      & = & \Any \cap (!X)^* \\
    \Through(X)    & = & \Out^* \cdot \In^* \cdot X \cdot \In^* \cdot \Out^* \\
    \Later(X)      & = & \Out^* \cdot (X \cap \Out) \cdot \Out^* \cdot \In^+ \cdot \Out^* \\
    \Before(X)     & = & \Out^* \cdot \In^+ \cdot \Out^* \cdot (X \cap \Out) \cdot \Out^* \\
    \End(X)        & = & \Any \cap (\Sigma^* \cdot X) \\
    \Start(X)      & = & \Any \cap (X \cdot \Sigma^*) \\
    \Exit(X)       & = & (\Out^* \cdot \In^* \cdot (X \cap \In) \cdot \Out \cdot \Out^*) \cup \\
                  &        & (\Out^* \cdot \In^+ \cdot (X \cap \Out) \cdot \Out^*) \\
    \Enter(X)      & = & (\Out^* \cdot \Out \cdot (X \cap \In) \cdot \In^* \cdot \Out^*) \cup \\
                  &        & (\Out^* \cdot (X \cap \Out) \cdot \In^+ \cdot \Out^*) \\
    \LinkKW(X,Y)     & = & \Any \cap (\Sigma^* \cdot X \cdot Y \cdot \Sigma^*) \\
    \PathKW(\vec{X}) & = & \Any \cap (\Sigma^* \cdot X_1 \dots X_n \cdot \Sigma^*) \\
    \Novalley(\vec{X}) & = & \Any ~ \cap \\
                  &   & \NOT\PathKW(X_2,X_1,X_2) ~ \cap \dots \cap \\
                  &   & \NOT\PathKW(X_n,X_{n-1},X_n) \\
  \end{array} \]

@@

\para{Syntax}
The \sysname FE is just a thin layer atop a core, regular-expression language (RIR) for describing preference-based path constraints. 
Figure~\ref{fig:rir-syntax} shows the RIR syntax. A policy consists of one or more constraints, each of which consists of a test on the type of route, and a corresponding set of preferred regular paths. Regular paths are regular expressions where the base characters are abstract locations---either a router or an AS. Special \In{} and \Out{} symbols refer to any internal or external location respectively.  In addition, $\Sigma$ refers to any symbol, and hence $\Sigma^*$ is
any sequence of symbols.  We also use other standard abbreviations
for regular expressions such as $r^+$, a sequence of one or more occurrences
of $r$, when appropriate.
%% For example, the constraint
%% \[(\textit{prefix}=74.3.28.0/24) \Link \\
%% (as200 \cdot \In^+) \Prefer (as100 \cdot \In^+)
%% \]
%% describes a more-preferred set of paths for traffic announced by a prefix no less specific than \CD{74.125.28.0/24}, which starts at AS 200, before entering and staying inside the user's network to get to the destination, and a less-preferred set of paths that start at AS 100 and are otherwise the same. The plus operator $\In^+$ stands for $\In \cdot \In^*$, at least one internal hop. Tests over route types use standard boolean connectives, and can refer to both prefixes and route community values.

@@

The semantics of RIR is in terms of ranked paths. Each preference-based regular path constraint (of the form $r_1 \Prefer \dots \Prefer r_j$) maps to a set of concrete paths in the network that match one of $r_i$. We denote a network path as a string of abstract locations (routers or external ASes) of the form: $n_1 n_2 \dots n_k$. A regular expression $r$ matches path $p$, if $p \in \mathcal{L}(r)$, that is the path is in the language of the regular expression, and $p$ is a topologically-valid path. We denote the length of a path $p$ as $\abs{p}$. A path $p$ will have a rank:
$$(\min_i \set{ p \in \mathcal{L}(r_i) }, \abs{p})$$
where the rank is lexicographically ordered according to (1) the most preferred regular expression matched, and (2) as a tie breaker, the length of the path. Lower ranks indicate \emph{more} preferred paths. Traffic may be sent along any of the most preferred paths for each pair of starting and ending locations that appear in some valid specified path.

@@

Intuitively, PGIR tracks which states of each automaton the policy is in as route announcements move between routers.
%
Consider the topology in Figure~\ref{fig:example-compilation}. Suppose we want a primary route from W that enters the network from $A$, and utilizes the A--C and C--D links. We also want a backup route that enters from $B$, and utilizes the B--C link, but is otherwise unconstrained. For simplicity, we assume that the route can end in either $X$, $Y$, or $Z$. The RIR for the policy is:
%
$$(\CD{W} \cdot \CD{A} \cdot \CD{C} \cdot \CD{D} \cdot \Out) \Prefer (\CD{W} \cdot \CD{B} \cdot \In^+ \cdot \Out)$$
%
Figure~\ref{fig:example-compilation} shows the policy automata for each regular expression preference. Since we are interested in the flow of control messages, the automata match backwards.
%
The figure also shows the PGIR after intersecting the topology and policy automata. Every path in it corresponds to a concrete path in the topology. In particular, every path through the PGIR that ends at a node $v$ such that the preference function $P(v) = \set{i_1, \dots, i_m}$ is non-empty, is a valid topological path that satisfies the policy constraints and results in a particular path with preferences $i_1$ through $i_m$.
%
For example, the path $X \cdot D \cdot C \cdot A \cdot W$ is a valid path in the topology that BGP route announcements might take, which would lead to obtaining the most prefered preference $1$.
BGP control messages can start from peer X, which would match the $\Out$ transition from both automata, leading to state $1$ in the first automaton, and state $1$ in the second automaton. This is reflected in the product graph by the node with state $(X,1,1)$. From here, if X were to advertise this route to D, it would result in the path $D X$, which would lead to state $2$ in the first automaton, and state $2$ in the second automaton, and so on.
%
The ``--" state indicates the corresponding automaton can not accept the current path or any extension of it. Since node $(W,5,-)$ is in an accepting state for the first automata, it indicates this path has preference 1.

@@

Finally, in our example, the external AS $W$ should prefer our internal router $A$ over $B$. In general, it is not possible to reliably control traffic entering the network beyond certain special cases. In this example, however, assuming our network and $W$ have an agreement to honor MEDs, the BGP MED attribute can influence $W$ to prefer $A$ over $B$. Additionally, the compiler can use the BGP no-export community to ensure that no other AS beyond $W$ can send us traffic. The compiler can perform a simple analysis to determine when it can utilize various BGP special attributes to ensure traffic enters the network in a particular way by looking at links in the product graph that cross from the internal topology to the external topology. 
%Aggregation is another tool commonly used to influence how traffic can enter the network, by relying on BGP's longest prefix matching default. Since aggregates are provided as additional user constraints, the compiler can also make use of this information to check that incoming preferences can be met given the aggregates specified (e.g., to ensure that all external ASes prefer to enter from one location over another).
Figure~\ref{fig:abgp-config} shows the full configuration from the compilation example.

After configuration generation, the compiler further processes the ABGP policy to minimize it, removing community tagging and matching when possible, combining filters, removing dead filters, and so on. In the compilation example, all community tags can be removed since there is never any ambiguity based on the router importing the route, and the neigbor the route is being imported from. Similarly, after removing the communities, the two rules at router $B$ are merged into a single rule.

>>>>

     pol     &::=& p_1, \dots, p_n & \textit{constraints} \\
     p       &::=& t \hspace{.3em} \Path \hspace{.3em} r_1 \Prefer \dots \Prefer r_m & \textit{preferences} \\
     x       &::=& a.b.c.d/n & \textit{prefix} \\
     c       &::=& \cdots & \textit{community value} \\
     t       &::=& \True & \textit{true} \\
         &\BNFALT& \NOT t & \textit{negation} \\
         &\BNFALT& t_1 \OR{} t_2 & \textit{disjunction} \\
         &\BNFALT& t_1 \AND{} t_2 & \textit{conjunction} \\
         &\BNFALT& \textit{prefix} = x & \textit{prefix test} \\
         &\BNFALT& \textit{comm} = c & \textit{community test} \\
     r       &::=& n & \textit{AS number} \\
         &\BNFALT& \emptyset & \textit{empty set} \\
         &\BNFALT& \In & \textit{internal loc} \\
         &\BNFALT& \Out & \textit{external loc} \\
         &\BNFALT& r_1 \cup r_2 & \textit{union} \\
         &\BNFALT& r_1 \cap r_2 & \textit{intersection} \\
         &\BNFALT& r_1 \cdot r_2 & \textit{concatenation} \\
         &\BNFALT& \NOT(r) & \textit{path negation} \\
         &\BNFALT& r^* & \textit{repetition} \\
     l       &::=& r_1 \rightarrow r_2 & \textit{link pairs} \\
     cc     &::=& agg(x, l) \BNFALT tag(c, t, l) & \textit{control constraints} \\
  \end{array} \]

  \end{minipage}
  %
  ~~
  \vrule
  ~~
  %
  \begin{minipage}[t]{.5\linewidth}\small
  \hdr{\large Propane Expansions}{}
  \vspace*{-1\baselineskip}
  %
  \[\begin{array}{rcl}
    \hline
    \Any           & = & \Out^* \cdot \In^+ \cdot \Out^* \\
    \None          & = & \emptyset \\
    \Internal      & = & \In^+ \\
    \Only(X)       & = & \Any \cap X^* \\
    \Never(X)      & = & \Any \cap (!X)^* \\
    \Through(X)    & = & \Out^* \cdot \In^* \cdot X \cdot \In^* \cdot \Out^* \\
    \Later(X)      & = & \Out^* \cdot (X \cap \Out) \cdot \Out^* \cdot \In^+ \cdot \Out^* \\
    \Before(X)     & = & \Out^* \cdot \In^+ \cdot \Out^* \cdot (X \cap \Out) \cdot \Out^* \\
    \End(X)        & = & \Any \cap (\Sigma^* \cdot X) \\
    \Start(X)      & = & \Any \cap (X \cdot \Sigma^*) \\
    \Exit(X)       & = & (\Out^* \cdot \In^* \cdot (X \cap \In) \cdot \Out \cdot \Out^*) \cup \\
                  &        & (\Out^* \cdot \In^+ \cdot (X \cap \Out) \cdot \Out^*) \\
    \Enter(X)      & = & (\Out^* \cdot \Out \cdot (X \cap \In) \cdot \In^* \cdot \Out^*) \cup \\
                  &        & (\Out^* \cdot (X \cap \Out) \cdot \In^+ \cdot \Out^*) \\
    \LinkKW(X,Y)     & = & \Any \cap (\Sigma^* \cdot X \cdot Y \cdot \Sigma^*) \\
    \PathKW(\vec{X}) & = & \Any \cap (\Sigma^* \cdot X_1 \dots X_n \cdot \Sigma^*) \\
    \Novalley(\vec{X}) & = & \Any ~ \cap \\
                  &   & \NOT\PathKW(X_2,X_1,X_2) ~ \cap \dots \cap \\
                  &   & \NOT\PathKW(X_n,X_{n-1},X_n) \\
  \end{array} \]

@@

\para{Syntax}
The \sysname FE is just a thin layer atop a core, regular-expression language (RIR) for describing preference-based path constraints. 
Figure~\ref{fig:rir-syntax} shows the RIR syntax. A policy consists of one or more constraints, each of which have a test on the type of route, and a corresponding set of preferred regular paths. Regular paths are regular expressions where the base characters are abstract locations---either a router or an AS. Special \In{} and \Out{} symbols refer to any internal or external location respectively. In addition, $\Sigma$ refers to any symbol. We also use other standard abbreviations
for regular expressions such as $r^+$, a sequence of one or more occurrences
of $r$, when appropriate.
%% For example, the constraint
%% \[(\textit{prefix}=74.3.28.0/24) \Link \\
%% (as200 \cdot \In^+) \Prefer (as100 \cdot \In^+)
%% \]
%% describes a more-preferred set of paths for traffic announced by a prefix no less specific than \CD{74.125.28.0/24}, which starts at AS 200, before entering and staying inside the user's network to get to the destination, and a less-preferred set of paths that start at AS 100 and are otherwise the same. The plus operator $\In^+$ stands for $\In \cdot \In^*$, at least one internal hop. Tests over route types use standard boolean connectives, and can refer to both prefixes and route community values.

@@

The semantics of RIR is in terms of ranked paths. Each preference-based regular path constraint (of the form $r_1 \Prefer \dots \Prefer r_j$) maps to a set of concrete paths in the network that match one of $r_i$. We denote a network path as a string of abstract locations (routers or external ASes) of the form: $n_1 n_2 \dots n_k$. A regular expression $r$ matches path $p$, if $p \in \mathcal{L}(r)$, that is the path is in the language of the regular expression, and $p$ is a topologically-valid path. We denote the length of a path $p$ as $\abs{p}$. A path $p$ will have a rank:
$$(\min_i \set{ p \in \mathcal{L}(r_i) }, \abs{p})$$
where the rank is lexicographically ordered according to (1) the most preferred regular expression matched, and (2) as a tie breaker, the path length. Lower ranks indicate \emph{more} preferred paths. Traffic may be sent on any of the most preferred paths for each pair of starting and ending locations that appear in some specified path. There is an implicit lowest preference to drop traffic: $\emptyset$ when no other route is possible.

@@

Intuitively, PGIR tracks which states of each automaton the policy is in as route announcements move between routers.
%
Consider the topology in Figure~\ref{fig:example-compilation}. Suppose we want a primary route from W that enters the network from $A$, and utilizes the A--C and C--D links. We also want a backup route that enters from $B$, and utilizes the B--C link, but is otherwise unconstrained. For simplicity, we assume that the route can end in either $X$, $Y$, or $Z$. The RIR for the policy is:
%
$$(\CD{W} \cdot \CD{A} \cdot \CD{C} \cdot \CD{D} \cdot \Out) \Prefer (\CD{W} \cdot \CD{B} \cdot \In^+ \cdot \Out)$$
%
Figure~\ref{fig:example-compilation} shows the policy automata for each regular expression preference. Since we are interested in the flow of control messages, the automata match backwards.
%
The figure also shows the PGIR after intersecting the topology and policy automata. Every path in it corresponds to a concrete path in the topology. In particular, every path through the PGIR that ends at a node $v$ such that the preference function $P(v) = \set{i_1, \dots, i_m}$ is non-empty, is a valid topological path that satisfies the policy constraints and results in a particular path with preferences $i_1$ through $i_m$.
%
For example, the path $X \cdot D \cdot C \cdot A \cdot W$ is a valid path in the topology that BGP route announcements might take, which would lead to obtaining the most prefered preference $1$.
BGP control messages can start from peer X, which would match the $\Out$ transition from both automata, leading to state $1$ in the first automaton, and state $1$ in the second automaton. This is reflected in the product graph by the node with state $(X,1,1)$. From here, if X were to advertise this route to D, it would result in the path $D \cdot X$, which would lead to state $2$ in the first automaton, and state $2$ in the second automaton, and so on.
%
The ``--" state indicates the corresponding automaton can not accept the current path or any extension of it. Since node $(W,5,-)$ is in an accepting state for the first automata, it indicates this path has preference 1.

@@

Finally, the external AS $W$ should prefer our internal router $A$ over $B$. In general, it is not possible to reliably control traffic entering the network beyond certain special cases. In this example, however, assuming our network and $W$ have an agreement to honor MEDs, the MED attribute can influence $W$ to prefer $A$ over $B$. Additionally, the compiler can use the BGP no-export community to ensure that no other AS beyond $W$ can send us traffic. The compiler can perform a simple analysis to determine when it can utilize BGP special attributes to ensure traffic enters the network in a particular way by looking at links in the product graph that cross from the internal topology to the external topology. 
%Aggregation is another tool commonly used to influence how traffic can enter the network, by relying on BGP's longest prefix matching default. Since aggregates are provided as additional user constraints, the compiler can also make use of this information to check that incoming preferences can be met given the aggregates specified (e.g., to ensure that all external ASes prefer to enter from one location over another).
Figure~\ref{fig:abgp-config} shows the full configuration from the compilation example.

After configuration generation, the compiler further processes the ABGP policy, removing community tags when possible, combining filters, removing dead filters, and so on. In the compilation example, all community tags can be removed since there is never any ambiguity based on the router importing the route, and the neigbor the route is being imported from. Similarly, after removing the communities, the two rules at router $B$ are merged into a single rule.

### 2016-01-30 paper/introduction.tex
final edits

To achieve the first goal, we borrow the idea of using regular
expressions to specify network paths from
recent high-level SDN languages such as FatTire~\cite{fattire},
Merlin~\cite{foster:merlin}, and
NetKAT~\cite{netkat}.  However, our design also contains several key
departures from existing languages.  The most important one is semantic:  the paths specified
by programmers can extend from outside the operator's network to inside
the network, across several devices internally, and then out again.  This design
choice allows users to specify preferences about both external and internal
routes in the exact same way.
%No SDN programming language provides similar facilities.
In addition, we augment the algebra
of regular expressions to directly support a notion of {\em preferences} and provide a semantics in terms of sets of ranked paths. The preferences indicate fail-over behaviors:  among all specified paths that are still available,
the system guarantees that the distributed implementation will always use the highest-ranked ones.
%We also support network operators by defining a
%set of common abbreviations for entering, leaving, and traversing
%the user network and we provide abstractions for managing common features %of
%control plane algorithms such as route aggregation.  Operators use familiar logical operations
%to build routes in a modular and compositional manner from these
%primitives.  Our abbreviations are translated in to our regular-expression-based
%intermediate representation and compiled from there.

>>>>

To achieve the first goal, we borrow the idea of using regular
expressions to specify network paths from
recent high-level SDN languages such as FatTire~\cite{fattire},
Merlin~\cite{foster:merlin}, and
NetKAT~\cite{netkat}.  However, our design also contains several key
departures from existing languages.  The most important one is semantic:  the paths specified 
can extend from outside the operator's network to inside
the network, across several devices internally, and then out again. This design
choice allows users to specify preferences about both external and internal
routes in the exact same way.
In addition, we augment the algebra
of regular expressions to directly support a notion of {\em preferences} and provide a semantics in terms of sets of ranked paths. The preferences indicate fail-over behaviors:  among all specified paths that are still available,
the system guarantees that the distributed implementation will always use the highest-ranked ones.
Although we target a distributed implementation, the language is more general and could potentially be used in an SDN context.

%We also support network operators by defining a
%set of common abbreviations for entering, leaving, and traversing
%the user network and we provide abstractions for managing common features %of
%control plane algorithms such as route aggregation.  Operators use familiar logical operations
%to build routes in a modular and compositional manner from these
%primitives.  Our abbreviations are translated in to our regular-expression-based
%intermediate representation and compiled from there.
